Highly Available Transactions: Virtues and Limitations

### 2013-04-29 writeup/conclusion.tex
smaller, smaller

\section{Conclusion and Future Work}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---``always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are not. This
offers a useful connection between storage semantics studied in the
database and distributed systems communities.

>>>>

\section{Conclusions and Future Work}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---``always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are not. This
offers a useful connection between storage semantics studied in the
database and distributed systems communities.

### 2013-04-29 writeup/hat-analysis.tex
smaller, smaller

\textbf{Repeatable Read} isolation is a confusing property. For example,
some IBM products use the term for fully serializable
isolation~\cite{hat-hotos}, while Gray~\cite{gray-isolation}, Berenson
et al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all
interpret Repeatable Read as providing serializability for all
operations except for predicate-based reads (and so Repeatable Read is
identical to serializability in a key-value store without
predicate-based access). In this latter model, the \textit{Phantom
  Problem}, whereby two successive predicate-based reads return
different data~\cite{gray-isolation}, is the only kind of
non-serializable behavior allowed. In this paper, we reserve the use
of ``Repeatable Read'' to refer to these definitions, which we will
soon show are not HAT-compliant.

@@

Transactional atomicity (TA) is core to ACID guarantees. Although, at
least by the ACID acronym, it is not an ``isolation'' property, TA
restricts transactions' ability to view the effects of partially
completed transactions. Under TA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter all effects of $T_i$ are observed by $T_2$. Together with
item cut isolation, TA prevents Read Skew anomalies (Berenson et al.'s
A5A~\cite{ansicritique}). As an example of TA, because $T_2$ has read
$T_1$'s write to $y$, $T_2$ must observe $b=c=1$ (or later versions
for each key):
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of
$a$. Notably, TA requires Read Committed isolation: observing all
effects of a transaction implicitly requires observing the final
(committed) effects of a transaction as well.

@@

\vspace{.5em}\noindent\textit{{Monotonic reads}} requires that, within
a session, subsequent reads to a given object ``never return any
previous values''; reads from each item progress according to a total
order. The ordering of reads should respect any externally observable
total ordering on transactions.

\vspace{.5em}\noindent\textit{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted by
the client. Any order on transactions should also be consistent with
any precedence that a global observer would see.

\vspace{.5em}\noindent\textit{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s).  Any order on transactions should
respect the reads-from order.\vspace{.5em}

@@

\vspace{.5em}\noindent\textit{{Read your writes}} requires
that whenever a client reads a given data item after updating it, the
read returns the updated value (or a value with a higher ID).

\vspace{.5em}\noindent\textit{{PRAM}} (Pipelined Random Access
Memory) provides the illusion of serializing each session's operations
and is the combination of monotonic reads, monotonic writes, and read
your writes~\cite{herlihy-art}.

\vspace{.5em}\noindent\textit{{Causal
    consistency}}~\cite{causalmemory} is the combination of all of the
session guarantees~\cite{sessiontocausal} (alternatively, PRAM with
writes-follow-reads) and is also referred to by Adya as PL-2L
isolation~\cite{adya}).\vspace{.5em}

>>>>

\textbf{Repeatable Read} isolation is a confusing property. For
example, some IBM products use the term for fully serializable
isolation~\cite{hat-hotos}, while Gray~\cite{gray-isolation}, Berenson
et al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all
interpret Repeatable Read as providing serializability for all
operations except for predicate-based reads (and so Repeatable Read is
identical to serializability in a key-value store without
predicate-based access). In models like Adya's PL-2.99, the
\textit{Phantom Problem}, whereby two successive predicate-based reads
return different data~\cite{gray-isolation}, is the only kind of
non-serializable behavior allowed. In this paper, we reserve the use
of ``Repeatable Read'' to refer to these definitions, which we will
soon show are not HAT-compliant.

@@

\textbf{Transactional atomicity (TA)} is core to ACID guarantees. Although, at
least by the ACID acronym, it is not an ``isolation'' property, TA
restricts transactions' ability to view the effects of partially
completed transactions. Under TA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter all effects of $T_i$ are observed by $T_2$. Together with
item cut isolation, TA prevents Read Skew anomalies (Berenson et al.'s
A5A~\cite{ansicritique}). As an example of TA, because $T_2$ has read
$T_1$'s write to $y$, $T_2$ must observe $b=c=1$ (or later versions
for each key):
\vspace{-.5em}
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of
$a$. Notably, TA requires Read Committed isolation: observing all
effects of a transaction implicitly requires observing the final
(committed) effects of a transaction as well.

@@

\vspace{.5em}\noindent\textbf{{Monotonic reads}} requires that, within
a session, subsequent reads to a given object ``never return any
previous values''; reads from each item progress according to a total
order. The ordering of reads should respect any externally observable
total ordering on transactions.

\vspace{.5em}\noindent\textbf{{Monotonic writes}} requires that each
session's writes become visible in the order they were submitted by
the client. Any order on transactions should also be consistent with
any precedence that a global observer would see.

\vspace{.5em}\noindent\textbf{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s).  Any order on transactions should
respect the reads-from order.\vspace{.5em}

@@

\vspace{.5em}\noindent\textbf{{Read your writes}} requires
that whenever a client reads a given data item after updating it, the
read returns the updated value (or a value with a higher ID).

\vspace{.5em}\noindent\textbf{{PRAM}} (Pipelined Random Access
Memory) provides the illusion of serializing each session's operations
and is the combination of monotonic reads, monotonic writes, and read
your writes~\cite{herlihy-art}.

\vspace{.5em}\noindent\textbf{{Causal
    consistency}}~\cite{causalmemory} is the combination of all of the
session guarantees~\cite{sessiontocausal} (alternatively, PRAM with
writes-follow-reads) and is also referred to by Adya as PL-2L
isolation~\cite{adya}).\vspace{.5em}

### 2013-04-29 writeup/intro.tex
smaller, smaller

Our investigation is based on both impossibility results and several
constructive proof-of-concept algorithms. For example, Snapshot
Isolation (SI) and Repeatable Reads (RR) are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we prove
is unavailable. However, Read Committed isolation, Transactional
Atomicity, and many of the weaker data consistency models from
database and distributed systems are achievable via novel algorithms
that rely on multi-versioning and limited client-side caching. For
several guarantees, such as causal consistency with phantom prevention
and ANSI Repeatable Read, we consider a modified form of high
availability in which clients ``stick to'' (i.e., have affinity with)
at least one server (a property which is often implicit in the
distributed systems literature~\cite{herlihy-art, eiger, cac} but
which requires explicit consideration in a client-server replicated
database context). This sticky availability is easily implementable
and widely employed~\cite{eiger, vogels-defs} but is more restrictive
than traditional definitions of high availability.

>>>>

Our investigation is based on both impossibility results and several
constructive proof-of-concept algorithms. For example, Snapshot
Isolation (SI) and Repeatable Reads (RR) are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we prove
is unavailable. However, Read Committed isolation, Transactional
Atomicity, and many of the weaker data consistency models from
database and distributed systems are achievable via novel algorithms
that rely on multi-versioning and limited client-side caching. For
several guarantees, such as causal consistency with phantom prevention
and ANSI Repeatable Read, we consider a modified form of high
availability in which clients ``stick to'' (i.e., have affinity with)
at least one server (a property which is often implicit in the
distributed systems literature~\cite{herlihy-art, eiger, cac} but
which requires explicit consideration in a client-server replicated
database context). This sticky availability is widely
employed~\cite{eiger, vogels-defs} but is less restrictive (and
therefore more easily achievable) than traditional high availability.
